The resurgence of the linear optics \sihui{quantum} interferometer --- recent advances \& applications

### 2019-01-03 arxiv/paper.tex
Added references

\documentclass[pra,singlecolumn,superscriptaddress,twoside,pra]{revtex4}

%\documentclass[times,final]{elsarticle}
%\usepackage{revip}
\usepackage{framed,multirow,threeparttable}
\usepackage{hyperref,mathrsfs,amsmath,amssymb,latexsym,array,hhline}
\newcommand{\bra}[1]{\langle#1|}
\newcommand{\ket}[1]{|#1\rangle}
\newcommand{\op}[2]{\hat{\textbf{#1}}_{#2}}
\newcommand{\dagop}[2]{\hat{\textbf{#1}}_{#2}^\dag}
\usepackage{graphicx}
\usepackage{subcaption}
\captionsetup{compatibility=false}

%\usepackage[colorlinks]{hyperref}
\usepackage[dvipsnames]{xcolor}

\newcommand{\sihui}[1]{{\color{red}{#1}}}
\newcommand{\peter}[1]{{\color{red}{#1}}}
\newcommand{\comment}[1]{{\color{YellowGreen}{#1}}}
\newcommand{\ah}{\hat{a}}
\newcommand{\adagh}{\hat{a}^\dag}

\definecolor{newcolor}{rgb}{.8,.349,.1}

@@

%
% Authors
%
\author{Si-Hui Tan}
\affiliation{Singapore University of Technology and Design, 20 Dover Drive, Singapore 138682}
\affiliation{Centre for Quantum Technologies, National University of Singapore, Block S15, 3 Science Drive 2, Singapore 117543, Singapore}
%
\author{Peter P.~Rohde}
\affiliation{Centre for Quantum Software \& Information (CQSI), Faculty 
of Engineering \& Information Technology, University of Technology 
Sydney, NSW 2007, Australia}

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=0.4\columnwidth]{reck}
\caption{Efficient decomposition of arbitrary linear optics 
networks into a sequence of beamsplitters and phase-shifters. This setup describes a generalized $n\times n$ multi-port Mach-Zehnder interferometer. Adapted from \cite{bib:Reck1994}.} \label{fig:Reck}	
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=0.8\columnwidth]{configMZI}
\caption{(a) A decomposition of arbitrary linear optics networks into a sequence of subunitaries, $D_i$, made up of Mach-Zehnder interferometers, and is a universal linear-optical processing unit. These subunitaries are controlled via their photon amplitudes ($\alpha_{i,j}$) and phases ($\phi_{i,j}$). (b) A state-of-art implementation of a universal linear-optical processing unit is shown. The processor has 6 inputs entering from the left, and 6 outputs leaving to the right, and is constructed from a cascade of 15 Mach-Zehnder interferometers. The classical control and processing unit sets the parameters for the interferometer, and actively cools it. Adapted from \cite{bib:Carolan15}.} \label{fig:Carolan}	
\end{figure}

@@

\begin{figure}
\centering
\begin{subfigure}{.5\columnwidth}
  \centering
  \includegraphics[width=.5\columnwidth]{cluster_state}
  \caption{}
  \label{fig:cluster}
\end{subfigure}%
\begin{subfigure}{.5\columnwidth}
  \centering
  \includegraphics[width=.65\columnwidth]{fusion}
  \caption{}
  \label{fig:fusion}
\end{subfigure}
\caption{(a) The representation of cluster states as a graph. Vertices represent qubits initialised into the $\ket{+}$ state, while edges represent the application on {\sc cphase} gates, the ordering of which is irrelevant. (b) Fusion gates for joining smaller cluster states into larger ones. Both require only a single polarising beamsplitter, and waveplates. The type-I and -II fusion gates consume 1 or 2 qubits (photons) respectively, creating an edge between the remaining graphs. The type-I gate consumes one fewer photon, but requires number-resolved detection on the detected mode. The type-II gate requires only on/off photodetection, but consumes an additional photon. Thus, the type-I gate can be employed to fuse two Bell pairs into a 3-qubit cluster state, whereas the type-II gate will only grow clusters when beginning with at least 3 qubits per cluster. }
\label{fig:test}
\end{figure}

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=0.3\columnwidth]{qscissors}
\caption{Figure showing optical state truncation for $d=1$. The input state is $\ket{\phi}$, and when the photodetectors $D_1$ and $D_2$ register one and zero photons respectively, the output state is $\ket{\phi}_{\rm out}=c_0\ket{0}+c_1\ket{1}$ with an appropriate normalization factor. Adapted from \cite{bib:Pegg98}.} \label{fig:qscissors}	
\end{figure}

@@

\begin{figure}
\centering
\begin{subfigure}{.5\columnwidth}
  \centering
  \includegraphics[width=0.7\columnwidth]{full_architecture}
  \caption{}
  \label{fig:FL_arch}
\end{subfigure}%
\begin{subfigure}{0.5\columnwidth}
  \centering
  \includegraphics[width=0.8\columnwidth]{PBS_TB}
  \caption{}
  \label{fig:PBS_TB}
\end{subfigure}
\caption{(a) A fibre-loop architecture for implementing arbitrary linear optics operations upon a time-bin-encoded pulse-train. (b) Mapping between two polarisation-encoded qubits undergoing a polarising beamsplitter (PBS) operation, and its equivalent representation using 4 time-bins in a pulse-train. The PBS completely reflects (transmits) the vertical (horizontal) polarisations. The evolution of the four logical basis states, and their respective outputs, are shown explicitly. Writing out this PBS transformation in matrix form yields a permutation. Taking this permutation and relabelling the modes, we obtain the time-bin transformation shown underneath -- a simple swap of two of the four time-bins.}
\end{figure}

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=0.4\columnwidth]{MZ}	
\caption{An optical switch for time-bin qubits that uses a Mach-Zehnder interferometer with a dynamic, classicallly-controlled phase-shift in one path.}
\label{figMZ}
\end{figure}

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=0.3\columnwidth]{BS_model}
\caption{The Boson-Sampling model. A string of $n$ single
photons is prepared in $m$ optical modes. They are evolved
via a passive interferometer $U$. Finally the photon-statistics
are sampled from the distribution $p(s_1,\ldots,s_m)$.
} \label{fig:BosonSampling}	
\end{figure}

@@

\begin{figure}
\centering
\begin{subfigure}{.5\columnwidth}
  \centering
  \includegraphics[width=0.4\columnwidth]{phE}
  \caption{}
   \label{fig:MZ}
\end{subfigure}%
\begin{subfigure}{0.5\columnwidth}
  \centering
  \includegraphics[width=0.75\columnwidth]{MORDOR}
  \caption{}
  \label{fig:MORDOR}
\end{subfigure}
\caption{(a) A schematic diagram of the Mach-Zehnder interferometer comprising of two 50/50 beamsplitters with a phase shift $\varphi$ in one of its two paths.	(b) Architecture of the quantum Fourier transform interferometer for metrology using single-photon states. QFT: Quantum Fourier transform circuit, and $\varphi_k=(k-1)\varphi$. Single photons are injected into the input modes on the left-hand-side, and a $N$-fold coincidence measurement is conducted at the output on the right-hand-side.}
\end{figure}

@@

\section*{Acknowledgments}
PPR is funded by an ARC Future Fellowship (project FT160100397). This research was supported in part by the Singapore National Research Foundation under NRF Award No. NRF-NRFF2013-01. ST acknowldeges support from the Air Force Office of Scientific Research under AOARD grant FA2386-15-1-4082.

>>>>

\documentclass[pra,singlecolumn,superscriptaddress]{revtex4}

%\documentclass[times,final]{elsarticle}
%\usepackage{revip}
\usepackage{framed,multirow,threeparttable}
\usepackage{mathrsfs,amsmath,amssymb,latexsym,array,hhline}
\newcommand{\bra}[1]{\langle#1|}
\newcommand{\ket}[1]{|#1\rangle}
\newcommand{\op}[2]{\hat{\textbf{#1}}_{#2}}
\newcommand{\dagop}[2]{\hat{\textbf{#1}}_{#2}^\dag}
\usepackage{graphicx}
\usepackage{subcaption}
\captionsetup{compatibility=false}

%\usepackage[colorlinks]{hyperref}

\newcommand{\ah}{\hat{a}}
\newcommand{\adagh}{\hat{a}^\dag}

@@

%
% Authors
%
\author{Si-Hui Tan}\email{sihui\_tan@sutd.edu.sg}
\affiliation{Singapore University of Technology and Design, 20 Dover Drive, Singapore 138682}
\affiliation{Centre for Quantum Technologies, National University of Singapore, Block S15, 3 Science Drive 2, Singapore 117543, Singapore}
%

\author{Peter P.~Rohde}
\affiliation{Centre for Quantum Software \& Information (CQSI), Faculty 
of Engineering \& Information Technology, University of Technology 
Sydney, NSW 2007, Australia}

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=0.4\columnwidth]{Reck.pdf}
\caption{Efficient decomposition of arbitrary linear optics 
networks into a sequence of beamsplitters and phase-shifters. This setup describes a generalized $n\times n$ multi-port Mach-Zehnder interferometer. Adapted from \cite{bib:Reck1994}.} \label{fig:Reck}	
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=0.8\columnwidth]{configMZI.pdf}
\caption{(a) A decomposition of arbitrary linear optics networks into a sequence of subunitaries, $D_i$, made up of Mach-Zehnder interferometers, and is a universal linear-optical processing unit. These subunitaries are controlled via their photon amplitudes ($\alpha_{i,j}$) and phases ($\phi_{i,j}$). (b) A state-of-art implementation of a universal linear-optical processing unit is shown. The processor has 6 inputs entering from the left, and 6 outputs leaving to the right, and is constructed from a cascade of 15 Mach-Zehnder interferometers. The classical control and processing unit sets the parameters for the interferometer, and actively cools it. Adapted from \cite{bib:Carolan15}.} \label{fig:Carolan}	
\end{figure}

@@

\begin{figure}
\centering
\begin{subfigure}{.5\columnwidth}
  \centering
  \includegraphics[width=.5\columnwidth]{cluster_state.pdf}
  \caption{}
  \label{fig:cluster}
\end{subfigure}%
\begin{subfigure}{.5\columnwidth}
  \centering
  \includegraphics[width=.65\columnwidth]{fusion.pdf}
  \caption{}
  \label{fig:fusion}
\end{subfigure}
\caption{(a) The representation of cluster states as a graph. Vertices represent qubits initialised into the $\ket{+}$ state, while edges represent the application on {\sc cphase} gates, the ordering of which is irrelevant. (b) Fusion gates for joining smaller cluster states into larger ones. Both require only a single polarising beamsplitter, and waveplates. The type-I and -II fusion gates consume 1 or 2 qubits (photons) respectively, creating an edge between the remaining graphs. The type-I gate consumes one fewer photon, but requires number-resolved detection on the detected mode. The type-II gate requires only on/off photodetection, but consumes an additional photon. Thus, the type-I gate can be employed to fuse two Bell pairs into a 3-qubit cluster state, whereas the type-II gate will only grow clusters when beginning with at least 3 qubits per cluster. }
\label{fig:test}
\end{figure}

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=0.3\columnwidth]{qscissors.pdf}
\caption{Figure showing optical state truncation for $d=1$. The input state is $\ket{\phi}$, and when the photodetectors $D_1$ and $D_2$ register one and zero photons respectively, the output state is $\ket{\phi}_{\rm out}=c_0\ket{0}+c_1\ket{1}$ with an appropriate normalization factor. Adapted from \cite{bib:Pegg98}.} \label{fig:qscissors}	
\end{figure}

@@

\begin{figure}
\centering
\begin{subfigure}{.5\columnwidth}
  \centering
  \includegraphics[width=0.7\columnwidth]{full_architecture.pdf}
  \caption{}
  \label{fig:FL_arch}
\end{subfigure}%
\begin{subfigure}{0.5\columnwidth}
  \centering
  \includegraphics[width=0.8\columnwidth]{PBS_TB.pdf}
  \caption{}
  \label{fig:PBS_TB}
\end{subfigure}
\caption{(a) A fibre-loop architecture for implementing arbitrary linear optics operations upon a time-bin-encoded pulse-train. (b) Mapping between two polarisation-encoded qubits undergoing a polarising beamsplitter (PBS) operation, and its equivalent representation using 4 time-bins in a pulse-train. The PBS completely reflects (transmits) the vertical (horizontal) polarisations. The evolution of the four logical basis states, and their respective outputs, are shown explicitly. Writing out this PBS transformation in matrix form yields a permutation. Taking this permutation and relabelling the modes, we obtain the time-bin transformation shown underneath -- a simple swap of two of the four time-bins.}
\end{figure}

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=0.4\columnwidth]{MZ.pdf}	
\caption{An optical switch for time-bin qubits that uses a Mach-Zehnder interferometer with a dynamic, classicallly-controlled phase-shift in one path.}
\label{figMZ}
\end{figure}

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=0.3\columnwidth]{BS_model.pdf}
\caption{The Boson-Sampling model. A string of $n$ single
photons is prepared in $m$ optical modes. They are evolved
via a passive interferometer $U$. Finally the photon-statistics
are sampled from the distribution $p(s_1,\ldots,s_m)$.
} \label{fig:BosonSampling}	
\end{figure}

@@

\begin{figure}
\centering
\begin{subfigure}{.5\columnwidth}
  \centering
  \includegraphics[width=0.4\columnwidth]{phE.pdf}
  \caption{}
   \label{fig:MZ}
\end{subfigure}%
\begin{subfigure}{0.5\columnwidth}
  \centering
  \includegraphics[width=0.75\columnwidth]{MORDOR.pdf}
  \caption{}
  \label{fig:MORDOR}
\end{subfigure}
\caption{(a) A schematic diagram of the Mach-Zehnder interferometer comprising of two 50/50 beamsplitters with a phase shift $\varphi$ in one of its two paths.	(b) Architecture of the quantum Fourier transform interferometer for metrology using single-photon states. QFT: Quantum Fourier transform circuit, and $\varphi_k=(k-1)\varphi$. Single photons are injected into the input modes on the left-hand-side, and a $N$-fold coincidence measurement is conducted at the output on the right-hand-side.}
\end{figure}

@@

\section*{Acknowledgments}
The authors thank A.~Laing and J.~O'Brien for their permission to use Figure 1 of \cite{bib:Carolan15}. PPR is funded by an ARC Future Fellowship (project FT160100397). This research was supported in part by the Singapore National Research Foundation under NRF Award No.~NRF-NRFF2013-01. ST acknowldeges support from the Air Force Office of Scientific Research under AOARD grant FA2386-15-1-4082.

### 2019-01-03 submitted_v2/review_v3/paper.tex
Added references

Alternatively, \sihui{Mach-Zehnder} interferometers can also be employed as building blocks instead of beamsplitters and phase-shifters \cite{bib:Englert2001}. \sihui{Such a decomposition and its implementation on an integrated photonic chip are shown in Fig.~\ref{fig:Carolan}. } Later, it was shown that any nontrivial beamsplitter, that does more than permuting modes or adding phases to them, is universal for linear optics \cite{bib:Bouland2014}. However, they do not provide an explicit construction for arbitrary unitaries.

@@

 As discussed in Section \ref{sec:bosonsampling}, the resources needed for Boson-sampling are readily available, but executes a task that is believed to be hard classically. Surprisingly, an entanglement between a large number of path labels is possible with single photons in a passive linear optical device. Motes {\it et al.~}harnessed this so-called number-path entanglement for quantum metrology without the use of entangled states \cite{bib:Motes15_LOQM,bib:Olsen2017}. Here, the phase shift to be estimated, $\varphi$, manifests itself as a linear phase gradient across $N$ modes (See Fig.~\ref{fig:MORDOR}). By straddling these modes among a quantum Fourier transform and its conjugate operation, it is possible to achieve $\Delta \varphi=\mathcal{O}(N^{-3/2})$ using a coincidence measurement with just a string of $N$ single photons, for up to at least $N=25$ modes. \sihui{Shortly after its construction, this scheme was realized experimentally \cite{bib:Zu-En2017}. The quantum Cram\'{e}r-Rao bound of estimating $\varphi$ in such interferometers was also analytically computed \cite{bib:Chenglong2017}.}

@@

Without the use of repeaters, it has been shown that the rate transmission of a channel with loss without entanglement assistance has a linear fallout \cite{bib:TGW14}, and its bound is known as the Takeoka-Guha-Wilde (TGW) bound. The minimal resources required to exceed the TGW bound is a subject of active research, and so far it has been shown that the performance of the repeater schemes of Ewert-Bergmann-van Loock \cite{bib:Ewert16}, and Sinclair et al. \cite{bib:Sinclair14} are able to exceed the TGW bound. Other benchmarking tools have also been used to analyze some of these repeater schemes in detail \cite{bib:Muralidhara16}. In terms of implementation, the greatest distance that a quantum state has been teleported to-date is 1400 km \cite{bib:Ren17}, and is demonstrated between an earth-based source with a detection unit on a satellite in space. This is a culmination of many decades of research from many groups in extending this distance. If it could be made to work with other basic units of quantum repeater networks that have been implemented \cite{bib:Kimble08,bib:Sangouard11, bib:MATN15}, a large-scale, long-distance repeater network could soon be possible.

}

@@

\section*{Acknowledgments}
PPR is funded by an ARC Future Fellowship (project FT160100397). This research was supported in part by the Singapore National Research Foundation under NRF Award No. NRF-NRFF2013-01. ST acknowldeges support from the Air Force Office of Scientific Research under AOARD grant FA2386-15-1-4082.

>>>>

Alternatively, \sihui{Mach-Zehnder} interferometers can also be employed as building blocks instead of beamsplitters and phase-shifters \cite{bib:Carolan15,bib:Englert2001, bib:Clements16}. \sihui{Such a decomposition and its implementation on an integrated photonic chip are shown in Fig.~\ref{fig:Carolan}. The scheme deployed by Clements {\it et al.}~\cite{bib:Clements16} requires only half the optical depth of Reck {\it et al.} and is robust to optical losses. A different route was taken by de Guise et al.~\cite{bib:deguise18} following a technique that decomposes a $n \times n$ unitary matrix into a beamsplitter sandwiched by two $(n-1) \times (n-1)$ unitary matrices. The technique can be used recursively to an elegant sequence of adjacent beamsplitters. This could be good experimentally for hard coded photonic circuits because it is difficult to mix non-adjacent modes there. The method of Clements et al.~share this advantage, but that of de Guise et al.~has an added advantage, that is, it uses fewer infinitesimal generators to get the same $n \times n$ unitary matrix. When the size of $n$ gets large and/or  the number of photons in the system is also large, the size of the full Hilbert space becomes large and a decomposition that uses fewer generators becomes obviously advantageous.} Later, it was shown by Bouland and Aaronson \cite{bib:Bouland2014} that any nontrivial beamsplitter, that does more than permuting modes or adding phases to them, is universal for linear optics. However, they do not provide an explicit construction for arbitrary unitaries.

@@

 As discussed in Section \ref{sec:bosonsampling}, the resources needed for Boson-sampling are readily available, but executes a task that is believed to be hard classically. Surprisingly, an entanglement between a large number of path labels is possible with single photons in a passive linear optical device. Motes {\it et al.~}harnessed this so-called number-path entanglement for quantum metrology without the use of entangled states \cite{bib:Motes15_LOQM,bib:Olsen2017}. Here, the phase shift to be estimated, $\varphi$, manifests itself as a linear phase gradient across $N$ modes (See Fig.~\ref{fig:MORDOR}). By straddling these modes among a quantum Fourier transform and its conjugate operation, it is possible to achieve $\Delta \varphi=\mathcal{O}(N^{-3/2})$ using a coincidence measurement with just a string of $N$ single photons, for up to at least $N=25$ modes. \sihui{Shortly after its construction, this scheme was realized experimentally \cite{bib:Zu-En2017}. The quantum Cram\'{e}r-Rao bound of estimating $\varphi$ in such interferometers was also analytically computed \cite{bib:Chenglong2017}. Also making use of entanglement, in this case, that of continuous variable multipartite entanglement, Zhuang et al.~showed that Heisenberg scaling in $N$ can be achieved when sensing a small phase shift that is distributed across $N$ modes \cite{ZZS18}.
 }

@@

%Without the use of repeaters, it has been shown that the rate transmission of a 
%channel with loss without entanglement assistance has a linear fallout 
%\cite{bib:TGW14}, and its bound is known as the Takeoka-Guha-Wilde (TGW) bound. 
%The minimal resources required to exceed the TGW bound is a subject of active 
%research, and so far it has been shown that the performance of the repeater 
%schemes of Ewert-Bergmann-van Loock \cite{bib:Ewert16}, and Sinclair et al. 
%\cite{bib:Sinclair14} are able to exceed the TGW bound. 
Without the use of repeaters it has been shown that the asymptotic rate of transmission of a channel with loss without entanglement assistance has a linear fallout according to the Takeoka-Guha-Wilde (TGW) bound \cite{bib:TGW14} and the tighter Pirandola-Laurenza-Ottaviani-Banchi (PLOB) bound \cite{PLOB17}. In particular, the PLOB bound coincides with the best known lower bound \cite{PGBL09} and establishes the two-way quantum capacity of the lossy channel. These results were refined by Wilde, Tomamichel and Berta who derived non-asymptotic rates that are applicable to protocols using such a channel finitely many times \cite{WTB17}. Others have generalized the TGW bound for arbitrary network topology \cite{AML16}. The minimal resources required to exceed the linear fallout is a subject of active research, and so far it has been beaten by the repeater schemes of Ewert-Bergmann-van Loock \cite{bib:Ewert16}, and Sinclair et al.~\cite{bib:Sinclair14}.

Other benchmarking tools have also been used to analyze some of these repeater schemes in detail \cite{bib:Muralidhara16}. In terms of implementation, the greatest distance that a quantum state has been teleported to-date is 1400 km \cite{bib:Ren17}, and is demonstrated between an earth-based source with a detection unit on a satellite in space. This is a culmination of many decades of research from many groups in extending this distance. If it could be made to work with other basic units of quantum repeater networks that have been implemented \cite{bib:Kimble08,bib:Sangouard11, bib:MATN15}, a large-scale, long-distance repeater network could soon be possible.
}

@@

\section*{Acknowledgments}
\sihui{We thank Hubert de Guise, Stefano Pirandola, Marco Tomamichel, and Quntao Zhuang for useful comments.} PPR is funded by an ARC Future Fellowship (project FT160100397). This research was supported in part by the Singapore National Research Foundation under NRF Award No.~NRF-NRFF2013-01. ST acknowledges support from the Air Force Office of Scientific Research under AOARD grant FA2386-15-1-4082 \sihui{and FA2386-18-1-4003. ST conducted part of this writing while she was a Guest Researcher at the Niels Bohr International Academy.}
